---
title: "Mapping Specifications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping Specifications}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Mapping Workflows

Currently, there exist 13 domains that have respective workflows that require mapping from source/raw data:

-   AE
-   COUNTRY
-   DATACHG
-   DATAENT
-   ENROLL
-   LB
-   PD
-   QUERY
-   SDRGCOMP
-   SITE
-   STUDCOMP
-   STUDY
-   SUBJ

Each of the above 13 domains generally map directly 1-to-1 from a raw source, except `COUNTRY` which relies on the same dataset as `SUBJ`. The following variables/types from their respective domains is **REQUIRED** in the source/raw data:

```{r raw, echo = FALSE, message = FALSE, error = FALSE}
library(dplyr)
library(DT)
library(gsm)
library(gsm.mapping)
# Source Data
lSource <- list(
    Source_SUBJ = clindata::rawplus_dm,
    Source_AE = clindata::rawplus_ae,
    Source_PD = clindata::ctms_protdev,
    Source_LB = clindata::rawplus_lb,
    Source_STUDCOMP = clindata::rawplus_studcomp,
    Source_SDRGCOMP = clindata::rawplus_sdrgcomp %>%
      dplyr::filter(.data$phase == 'Blinded Study Drug Completion'),
    Source_DATACHG = clindata::edc_data_points,
    Source_DATAENT = clindata::edc_data_pages,
    Source_QUERY = clindata::edc_queries,
    Source_ENROLL = clindata::rawplus_enroll,
    Source_SITE = clindata::ctms_site,
    Source_STUDY = clindata::ctms_study
)

# Step 0 - Data Ingestion - standardize tables/columns names
lRaw <- list(
    Raw_SUBJ = lSource$Source_SUBJ,
    Raw_AE = lSource$Source_AE,
    Raw_PD = lSource$Source_PD %>%
      rename(subjid = subjectenrollmentnumber),
    Raw_LB = lSource$Source_LB,
    Raw_STUDCOMP = lSource$Source_STUDCOMP %>%
      select(subjid, compyn),
    Raw_SDRGCOMP = lSource$Source_SDRGCOMP,
    Raw_DATACHG = lSource$Source_DATACHG %>%
      rename(subject_nsv = subjectname),
    Raw_DATAENT = lSource$Source_DATAENT %>%
      rename(subject_nsv = subjectname),
    Raw_QUERY = lSource$Source_QUERY %>%
      rename(subject_nsv = subjectname),
    Raw_ENROLL = lSource$Source_ENROLL,
    Raw_SITE = lSource$Source_SITE %>%
      rename(studyid = protocol) %>%
      rename(invid = pi_number) %>%
      rename(InvestigatorFirstName = pi_first_name) %>%
      rename(InvestigatorLastName = pi_last_name) %>%
      rename(City = city) %>%
      rename(State = state) %>%
      rename(Country = country) %>%
      rename(Status = site_status),
    Raw_STUDY = lSource$Source_STUDY %>%
      rename(studyid = protocol_number) %>%
      rename(Status = status)
)

# Step 1 - Create Mapped Data Layer - filter, aggregate and join raw data to create mapped data layer
mappings_wf <- MakeWorkflowList(strPath = "workflow/1_mappings")

# Extract the nested variables and their types correctly
raw_result <- lapply(names(mappings_wf), function(domain_name) {
  domain <- mappings_wf[[domain_name]]$spec
  do.call(rbind, lapply(names(domain), function(var_name) {
    variables <- domain[[var_name]]
    data.frame(
      Domain = domain_name,
      Variable = names(variables),
      Type = sapply(variables, function(var) var$type),
      stringsAsFactors = FALSE
    )
  }))
})
# Combine all domains into one table
raw_final_table <- do.call(rbind, raw_result)
row.names(raw_final_table) <- NULL
raw_final_table %>%
    DT::datatable(
      extensions = 'FixedColumns',
      options = list(
        scrollX = FALSE,
        fixedColumns = TRUE
      ),
      rownames = FALSE
    )
```

```{r mapped, eval = FALSE, echo = FALSE, message = FALSE}
mapped <- RunWorkflows(mappings_wf, lRaw)

raw_names <- lapply(lRaw, function(x) names(x))
raw_names2 <- raw_names[sort(names(raw_names))]
mapped_names <- lapply(mapped, function(x) names(x))
mapped_names2 <- mapped_names[sort(names(mapped_names))]
mapped_names2 <- mapped_names2[-which(names(mapped_names2) == "Mapped_COUNTRY")]
map2(.x = raw_names2, .y = mapped_names2, .f = \(raw_names2, mapped_names2) mapped_names2[-(which(mapped_names2 %in% raw_names2))])
```
