---
title: "Mapping Specifications"
description: "A Vignette that details all mapping specifications for all domains in {gsm.mapping}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping Specifications}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Mapping Workflows

```{r discover_domains, echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
library(DT)
library(gsm.core)
library(gsm.mapping)
library(dplyr)
library(reactable)
library(yaml)
library(htmltools)

# Automatically discover all mapping domains from the inst/workflow/1_mappings folder
mapping_files <- list.files(
  system.file("workflow/1_mappings", package = "gsm.mapping"), 
  pattern = "\\.yaml$", 
  full.names = TRUE
)
all_domains <- tools::file_path_sans_ext(basename(mapping_files))
```

Currently, there exist **`r length(all_domains)` domains** that have mapping specifications for transforming raw/source data into standardized formats for GSM workflows. Each mapping specification defines the required variables, their data types, and transformation rules needed to create consistent, analysis-ready datasets.

The table below shows all available domain mappings with their specifications. Use the domain dropdown filter to select specific domains, or use the search and filter functionality to find specific variable types. Hover over truncated descriptions and dataset names to see the full text.

### All Domain Mappings
```{r all_mappings, echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
# Create Mapped Data Layer for all discovered domains
mappings_wf <- MakeWorkflowList(strNames = all_domains, strPath = "workflow/1_mappings", strPackage = "gsm.mapping")

# Extract domain metadata from YAML files
domain_metadata <- lapply(mapping_files, function(file) {
  yaml_content <- yaml::read_yaml(file)
  domain_name <- tools::file_path_sans_ext(basename(file))
  
  list(
    Domain = domain_name,
    Description = ifelse(!is.null(yaml_content$meta$Description), 
                         yaml_content$meta$Description, 
                         "No description available")
  )
})

# Convert metadata to data frame
metadata_df <- do.call(rbind, lapply(domain_metadata, data.frame, stringsAsFactors = FALSE))

# Extract the nested variables and their types correctly
raw_result <- lapply(names(mappings_wf), function(domain_name) {
  domain <- mappings_wf[[domain_name]]$spec
  do.call(rbind, lapply(names(domain), function(var_name) {
    variables <- domain[[var_name]]
    data.frame(
      Domain = domain_name,
      Dataset = var_name,
      Source_Variable = sapply(seq_along(variables), function(i){
        ifelse(is.null(variables[[i]]$source_col), 
               names(variables)[i], 
               variables[[i]]$source_col)
        }),
      Mapped_Variable = names(variables),
      Type = sapply(variables, function(var) var$type),
      stringsAsFactors = FALSE
    )
  }))
})

# Combine all domains into one table
raw_final_table <- do.call(rbind, raw_result)
row.names(raw_final_table) <- NULL

# Merge with metadata
final_table <- merge(raw_final_table, metadata_df, by = "Domain", all.x = TRUE)

# Reorder columns for better presentation
final_table <- final_table %>%
  dplyr::select(Domain, Description, Dataset, Source_Variable, Mapped_Variable, Type) %>%
  dplyr::arrange(Domain)


# Create the table with an elementId for JavaScript filtering
table_widget <- reactable::reactable(
  final_table,
  elementId = "domain-mapping-table",
  defaultPageSize = 50,
  searchable = TRUE,
  filterable = TRUE,
  sortable = TRUE,
  resizable = TRUE,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(25, 50, 100, 200),
  bordered = FALSE,
  highlight = TRUE,
  striped = TRUE,
  compact = TRUE,
  wrap = FALSE,
  theme = reactable::reactableTheme(
    borderColor = "#dee2e6",
    backgroundColor = "white",
    stripedColor = "#f8f9fa",
    style = list(
      fontFamily = "Arial, sans-serif",
      fontSize = "12px"
    ),
    headerStyle = list(
      fontWeight = "bold",
      fontSize = "13px",
      backgroundColor = "#e9ecef",
      borderBottom = "2px solid #6c757d",
      color = "#495057"
    ),
    groupHeaderStyle = list(
      fontWeight = "bold",
      backgroundColor = "#f1f3f4",
      borderBottom = "1px solid #dee2e6"
    )
  ),
  columns = list(
    Domain = reactable::colDef(
      name = "Domain", 
      style = list(fontWeight = "bold", fontSize = "13px"),
      minWidth = 160,
      filterInput = function(values, name) {
        tags$select(
          # Add "All" option at the beginning
          tags$option(value = "", "All Domains"),
          # Add unique domain options
          lapply(unique(sort(values)), function(value) {
            tags$option(value = value, value)
          }),
          onchange = sprintf("Reactable.setFilter('%s', '%s', this.value)", 
                           "domain-mapping-table", name)
        )
      }
    ),
    Description = reactable::colDef(
      name = "Description",
      style = list(fontSize = "11px", color = "#6c757d"),
      cell = function(value) {
        div(title = value, 
            style = "cursor: help; overflow: hidden; text-overflow: ellipsis; white-space: nowrap;", 
            value)
      }
    ),
    Dataset = reactable::colDef(
      name = "Dataset", 
      minWidth = 100,
      style = list(fontSize = "12px"),
      cell = function(value) {
        div(title = value, 
            style = "cursor: help; overflow: hidden; text-overflow: ellipsis; white-space: nowrap;", 
            value)
      }
    ),
    Source_Variable = reactable::colDef(
      name = "Source Variable", 
      style = list(fontSize = "12px", fontFamily = "monospace")
    ),
    Mapped_Variable = reactable::colDef(
      name = "Mapped Variable",
      style = list(fontSize = "12px", fontFamily = "monospace")
    ),
    Type = reactable::colDef(
      name = "Type", 
      width = 90,
      style = function(value) {
        color <- switch(tolower(value),
                       "character" = "#007bff",
                       "integer" = "#28a745",
                       "numeric" = "#28a745", 
                       "date" = "#dc3545",
                       "timestamp" = "#6f42c1",
                       "#6c757d")
        list(color = color, fontWeight = "500", fontSize = "11px")
      }
    )
  ),
  defaultColDef = reactable::colDef(
    style = list(fontSize = "12px"),
    headerStyle = list(fontSize = "12px")
  )
)

# Display the table
table_widget
```
