---
title: "Mapping Specifications"
description: "A Vignette that details all mapping specifications for all domains in {gsm.mapping}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping Specifications}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Mapping Workflows

Currently, there exist 19 domains that have respective workflows that require mapping from source/raw data. 13 are general domains used across several packages and extensions:

-   AE
-   DATACHG
-   DATAENT
-   ENROLL
-   LB
-   PD
-   PK
-   QUERY
-   SDRGCOMP
-   SITE
-   STUDCOMP
-   STUDY
-   SUBJ

And 6 are domains that are currently specific to the `gsm.endpoints` extension package:

- Baseline
- Consents
- Death
- OverallResponse
- Randomization
- Visit

Each of the above 20 domains generally map directly 1-to-1 from a raw source, except `COUNTRY` which relies on the same dataset as `SUBJ`. The following variables/types from their respective domains is **REQUIRED** in the source/raw data:

### Core Mappings
```{r core, echo = FALSE, message = FALSE, error = FALSE}
library(DT)
library(gsm.core)
library(gsm.mapping)
library(dplyr)
library(reactable)

core_mappings <- c("AE", "COUNTRY", "DATACHG", "DATAENT", "ENROLL", "LB",
                   "PD", "QUERY", "STUDY", "STUDCOMP", "SDRGCOMP", "SITE", "SUBJ")

# Step 1 - Create Mapped Data Layer - filter, aggregate and join raw data to create mapped data layer
mappings_wf <- MakeWorkflowList(strNames = core_mappings, strPath = "workflow/1_mappings", strPackage = "gsm.mapping")

# Extract the nested variables and their types correctly
raw_result <- lapply(names(mappings_wf), function(domain_name) {
  domain <- mappings_wf[[domain_name]]$spec
  do.call(rbind, lapply(names(domain), function(var_name) {
    variables <- domain[[var_name]]
    data.frame(
      Domain = domain_name,
      Source_Variable = sapply(seq_along(variables), function(i){
        ifelse(is.null(variables[[i]]$source_col), 
               names(variables)[i], 
               variables[[i]]$source_col)
        }),
      Mapped_Variable = names(variables),
      Type = sapply(variables, function(var) var$type),
      stringsAsFactors = FALSE
    )
  }))
})
# Combine all domains into one table
raw_final_table <- do.call(rbind, raw_result)
row.names(raw_final_table) <- NULL

reactable(
  raw_final_table,
  groupBy = "Domain",
  defaultPageSize = 15,
  searchable = TRUE,
  bordered = FALSE, # Removes unnecessary borders
  highlight = TRUE, # Highlights row on hover
  theme = reactableTheme(
    borderColor = "#dee2e6",
    backgroundColor = "white",
    style = list(
      fontFamily = "Arial, sans-serif",
      fontSize = "14px"
    ),
    headerStyle = list(
      fontWeight = "bold",
      fontSize = "16px",
      backgroundColor = "#f8f9fa",
      borderBottom = "2px solid #dee2e6"
    )
  ),
  columns = list(
    Domain = colDef(name = "Domain", style = list(fontWeight = "bold", fontSize = "15px")),
    Source_Variable = colDef(name = "Source Variable"),
    Mapped_Variable = colDef(name = "Mapped Variable"),
    Type = colDef(name = "Type")
  )
)
```


## Endpoint Mappings

```{r endpoints, echo = FALSE, message = FALSE, error = FALSE}
endpoint_mappings <- c("Baseline", "Consents", "Death", "OverallResponse", "Randomization",
                   "Visit", "STUDY", "STUDCOMP","SITE", "SUBJ")


# Step 1 - Create Mapped Data Layer - filter, aggregate and join raw data to create mapped data layer
mappings_ep <- MakeWorkflowList(strNames = endpoint_mappings, strPath = "workflow/1_mappings", strPackage = "gsm.mapping")

# Extract the nested variables and their types correctly
raw_result_ep <- lapply(names(mappings_ep), function(domain_name) {
  domain <- mappings_ep[[domain_name]]$spec
  do.call(rbind, lapply(names(domain), function(var_name) {
    variables <- domain[[var_name]]
    data.frame(
      Domain = domain_name,
      Source_Variable = sapply(seq_along(variables), function(i){
        ifelse(is.null(variables[[i]]$source_col), 
               names(variables)[i], 
               variables[[i]]$source_col)
        }),
      Mapped_Variable = names(variables),
      Type = sapply(variables, function(var) var$type),
      stringsAsFactors = FALSE
    )
  }))
})
# Combine all domains into one table
raw_final_table_ep <- do.call(rbind, raw_result_ep)
row.names(raw_final_table_ep) <- NULL
reactable(
  raw_final_table_ep,
  groupBy = "Domain",
  defaultPageSize = 15,
  searchable = TRUE,
  bordered = FALSE, # Removes unnecessary borders
  highlight = TRUE, # Highlights row on hover
  theme = reactableTheme(
    borderColor = "#dee2e6",
    backgroundColor = "white",
    style = list(
      fontFamily = "Arial, sans-serif",
      fontSize = "14px"
    ),
    headerStyle = list(
      fontWeight = "bold",
      fontSize = "16px",
      backgroundColor = "#f8f9fa",
      borderBottom = "2px solid #dee2e6"
    )
  ),
  columns = list(
    Domain = colDef(name = "Domain", style = list(fontWeight = "bold", fontSize = "15px")),
    Source_Variable = colDef(name = "Source Variable"),
    Mapped_Variable = colDef(name = "Mapped Variable"),
    Type = colDef(name = "Type")
  )
)
```
