---
title: "Example Mapping Workflow"
author: "[gsm.mapping](https://gilead-biostats.github.io/gsm.mapping) Example"
description: "Example of creating the Mapped Data Layer step-by-step using the standard mappings in gsm.mapping"
index: 1
date: "`r format(Sys.time(), '%B %d, %Y %H:%M:%S %Z')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Set Up

This page walks through creating the **Mapped Data Layer** step-by-step using the
standard mappings shipped with the `gsm.mapping` package. The Mapped Data Layer
filters, aggregates, and joins raw/source data into the standardized datasets that
downstream `gsm` packages (e.g. `gsm.kri`) expect.

Expand the "Setup" section below to see the libraries and source data used, or
scroll down to follow the workflow.

<details>
<summary>Setup</summary>

```{r}
# Load libraries
library(gsm.core)
library(gsm.mapping)
library(dplyr)

# The 12 standard domains used to build a KRI report
core_mappings <- c(
  "AE", "COUNTRY", "DATACHG", "DATAENT", "ENROLL", "LB",
  "PD", "QUERY", "STUDY", "STUDCOMP", "SDRGCOMP", "SITE", "SUBJ"
)

# Sample raw/source data provided by gsm.core
lSource <- gsm.core::lSource
```

</details>

# Create the Mapped Data Layer

The Mapped Data Layer is created in a short, four-step workflow: build the list of
mapping workflows, combine their specifications, ingest the raw data, and run the
workflows to produce the mapped data.

## Step 1 - Build the mapping workflows

`workr::MakeWorkflowList()` reads the YAML mapping specifications shipped in
`inst/workflow/1_mappings` for each requested domain.

```{r}
mappings_wf <- workr::MakeWorkflowList(
  strNames = core_mappings,
  strPath = "workflow/1_mappings",
  strPackage = "gsm.mapping"
)
```

## Step 2 - Combine the specifications

`CombineSpecs()` collapses the individual domain specifications into a single
specification describing every raw dataset and column required.

```{r}
mappings_spec <- CombineSpecs(mappings_wf)
```

## Step 3 - Ingest the raw data

`Ingest()` uses the combined specification to select and standardize the required
columns from the raw/source data.

```{r}
lRaw <- Ingest(lSource, mappings_spec)
```

## Step 4 - Run the workflows

`workr::RunWorkflows()` executes each mapping workflow against the ingested data to
filter, aggregate, and join it into the final Mapped Data Layer.

```{r}
mapped <- workr::RunWorkflows(mappings_wf, lRaw)
```

# Review the Output

`mapped` is a named list of standardized datasets, one per mapped domain, ready for
use in downstream `gsm` reporting workflows.

```{r}
names(mapped)
```
